Online Appendix

### 2021-04-19 paper/v2_taylor_francis/broadband.speed.covid.Rmd
submission

<!-- # Acknowledgement(s) {-} -->

<!-- An unnumbered section, e.g.\ `\section*{Acknowledgements}`, may be used for thanks, etc.\ if required and included _in the non-anonymous version_ before any Notes or References. -->

>>>>

# Supplemental Material {-}

The LAD upload clusters as well as the mean AM and PM upload speeds are presented 
in the Supplemental Material. Download the .html file and open it using your internet
browser. Hovering over the LAD polygons provides the mean internet speeds.

# Acknowledgement {-}

The authors would like to thank [https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)
for providing the main data used for this paper.

### 2021-04-19 paper/v2_taylor_francis/broadband.speed.covid.tex
submission

The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest
in different geographies in the UK
\citep{riddlesden2014broadband, philip2017digital}. Second level digital
divides consider the presence or lack of the necessary skills to
effectively utilise digital technologies and the internet
\citep{blank2014dimensions, van2011internet}. The third level focuses on
the heterogenous returns of internet usage among different socioeconomic
groups and, consequently, how digital technologies can assist in
bridging or further enhancing existing socioeconomic divides.
\citep{stern2009levels, van2014digital, van2015third}. The capability to
telecommute is related to all three levels of digital divides, but more
importantly leads to differentiated outcomes regarding the economic
resilience of people and places to overcome a systemic shock such as the
current pandemic.

Our understanding of telecommuting as a product of enabled occupations
can be described as a manifestation of the third level digital divide,
as those who are able to use digital technologies to work from home
benefit from a high rate of return on their use of the internet in terms
of autonomy, flexibility, and time saved from commuting
\citep{peters2004employees, siha2006telecommuting, singh2013modeling}.
In some aspects, these returns may have increased during the Covid-19
crisis, when those with the capability to telecommute also have the
ability to maintain their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the
first level digital divide, which is associated with access to and
quality of internet connectivity. \citet{SALEMINK2017360} provides a
systematic review of the pre-pandemic, first level digital divide in
infrastructure quality between urban and rural areas in various advanced
economies. Rural areas, predictably, fare worse. This finding broadly
applies in the UK, where studies that also analysed broadband speed
checks concluded that average speeds are lower in rural areas, something
that has not been improved by policy measures to increase competition
\citep{riddlesden2014broadband, nardotto2015unbundling}. In contrast,
\citet{riddlesden2014broadband} found that levels of deprivation did not
correlate with first level digital divides.

However, as \citet{blank2018local} highlight, variation in individual
internet availability and uptake is a product of more complex spatial
and demographic characteristics than simple rurality or urbanisation.
Dense urban areas were shown to suffer more from slowdown during peak
hours, although these services were more likely to be improved by
increased competition between providers, such as between new entrants
and Virgin Media cable connections
\citep{riddlesden2014broadband, nardotto2015unbundling}. The latter were
historically available to only \(45\)\% of premises in the UK
\citep{ofcom2016}, where the more lucrative and competitive market
originally attracted the cable TV provider. Whether the variation in
infrastructure quality and reliability affects the spatial footprint of
telecommuting has not previously been investigated, in part because
telecommuting has not previously been a cause of concentrated demand and
pressure on internet services.

Other intersections between digital divides have been subject to study.
There are indications that those who purchase high speed connections
consume more data of all sorts and use their connections for a greater
variety of purposes \citep{hauge2011consumer}. There is also a
correlation between access to internet services and a reduction in
household transport spend \citep{bris2017ict}. These studies suggest
that better internet services enable households to make savings and
efficiencies, an example of the first level digital divide reinforcing
the third level. Such digital intersections and efficiencies have also
changed transportation, connectivity and the structure of cities during
the pandemic. We now have enough data to confirm the drastic alteration
observed in the space-time geography of cities around the world in the
last year \citep[e.g.~][]{google2020, shibayama2021impact}. Yet,
although there is a broad agreement that these changes during the
pandemic have played a pivotal role in stopping the spread of the virus
\citep{jia2020population, yang2020taking, mu2020interplay}, the extent
to which the increased levels of working from home, the consequent
decrease in commuting flows, and the altered structure of cities will
remain post-pandemic is the subject of considerable debate in the
literature.

@@

Therefore, home-based digital infrastructure which considers upload
speeds and working day reliability as well as availability are likely to
be particularly important in a future where telecommuting might be a
more common means of accessing work and broadband services must be fit
for purpose. Although the long-term effects of such drastic changes in
telecommuting and attitudes towards working from home are difficult to
predict, the reliability of home broadband services deserve more
consideration than in the past. This would represent a switch from
previous demand-side broadband policies in the UK which tended to be
aimed at supporting small and medium enterprises
\citep{HENDERSON2020102024}. SUch policies were based on previous
research regarding the productivity effects of broadband infrastructure
\citep{DESTEFANO2018110}. However, this stream of research tended not to
consider residential locations as places that host economic activity.
Such policies also made some assumptions about the quality of service in
residential markets that could be assured without government
intervention -- an assumption which our analysis suggest is not entirely
accurate.

@@

>>>>

The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest
in different geographies in the UK
\citep{riddlesden2014broadband, philip2017digital}. Second level digital
divides consider the presence or lack of the necessary skills to
effectively utilise digital technologies and the internet
\citep{blank2014dimensions, van2011internet}. The third level focuses on
the heterogenous returns of internet usage among different socioeconomic
groups and, consequently, how digital technologies can assist in
bridging or further enhancing existing socioeconomic divides.
\citep{stern2009levels, van2014digital, van2015third}. The potential for
these different levels of digital divide to interact is growing in
parallel with the digitisation of various sectors of the economy and
society. For example, \citet{li2020challenges} highlights how the
intersection of new demands upon ICT infrastructure and a paucity of end
users means that some areas are being left behind in the roll-out of
smart applications and services. They argue that this could create a
geospatial `smart divide', with areas unable to capitalise on the
opportunities offered by such applications and services. Likewise, the
capability to telecommute is related to all three levels of digital
divides, but more importantly leads to differentiated outcomes regarding
the economic resilience of people and places to overcome a systemic
shock such as the current pandemic.

Our understanding of telecommuting as a product of enabled occupations
can be described as a manifestation of the third level digital divide,
as those who are able to use digital technologies to work from home
benefit from a high rate of return on their use of the internet in terms
of autonomy, flexibility, and time saved from commuting
\citep{peters2004employees, siha2006telecommuting, singh2013modeling}.
In some aspects, these returns may have increased during the Covid-19
crisis, when those with the capability to telecommute also have the
ability to maintain their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the
first level digital divide, which is associated with geospatial access
to and quality of internet connectivity. \citet{SALEMINK2017360}
provides a systematic review of the pre-pandemic, first level digital
divide in infrastructure quality between urban and rural areas in
various advanced economies. Rural areas, predictably, fare worse. This
finding broadly applies in the UK, where studies that also analysed
broadband speed checks concluded that average speeds are lower in rural
areas, something that has not been improved by policy measures to
increase competition
\citep{riddlesden2014broadband, nardotto2015unbundling}. In contrast,
\citet{riddlesden2014broadband} found that levels of deprivation did not
correlate with first level digital divides, perhaps because many
deprived areas benefit from spillover effects of quality installed
infrastructure nearby.

However, as \citet{blank2018local} highlight, variation in individual
internet availability and uptake is a product of more complex geospatial
and demographic characteristics than simple rurality, urbanisation or
levels of deprivation. Dense urban areas were shown to suffer more from
slowdown during peak hours, although these services were more likely to
be improved by increased competition between providers, such as between
new entrants and Virgin Media cable connections
\citep{riddlesden2014broadband, nardotto2015unbundling}. The latter were
historically available to only \(45\)\% of premises in the UK
\citep{ofcom2016}, where the more lucrative and competitive market
originally attracted the cable TV provider. Whether the geospatial
variation in infrastructure quality and reliability affects the spatial
footprint of telecommuting has not previously been investigated, in part
because telecommuting has not previously been a cause of concentrated
demand and pressure on internet services.

Other intersections between digital divides have been subject to study.
There are indications that those who purchase high speed connections
consume more data of all sorts and use their connections for a greater
variety of purposes \citep{hauge2011consumer}. There is also a
correlation between access to internet services and a reduction in
household transport spend \citep{bris2017ict}. These studies suggest
that better internet services enable households to make savings and
efficiencies, an example of the first level digital divide reinforcing
the third level. Such digital intersections and efficiencies have also
changed transportation, connectivity and the structure of cities during
the pandemic. We now have enough data to confirm the drastic alteration
observed in the space-time geography of cities around the world in the
last year \citep[e.g.][]{google2020, shibayama2021impact}. Yet, although
there is a broad agreement that these changes during the pandemic have
played a pivotal role in stopping the spread of the virus
\citep{jia2020population, yang2020taking, mu2020interplay}, the extent
to which the increased levels of working from home, the consequent
decrease in commuting flows, and the altered structure of cities will
remain post-pandemic is the subject of considerable debate in the
literature.

@@

Therefore, home-based digital infrastructure which considers upload
speeds and working day reliability as well as availability are likely to
be particularly important in a future where telecommuting might be a
more common means of accessing work and broadband services must be fit
for purpose. Although the long-term effects of such drastic changes in
telecommuting and attitudes towards working from home are difficult to
predict, the reliability of home broadband services deserve more
consideration than in the past. This would represent a switch from
previous demand-side broadband policies in the UK which tended to be
aimed at supporting small and medium enterprises
\citep{HENDERSON2020102024}. Such policies were based on previous
research regarding the productivity effects of broadband infrastructure
\citep{DESTEFANO2018110}. However, this stream of research tended not to
consider residential locations as places that host economic activity.
Such policies also made some assumptions about the quality of service in
residential markets that could be assured without government
intervention -- an assumption which our analysis suggest is not entirely
accurate.

@@

\hypertarget{supplemental-material}{%
\section*{Supplemental Material}\label{supplemental-material}}
\addcontentsline{toc}{section}{Supplemental Material}

The LAD upload clusters as well as the mean AM and PM upload speeds are
presented in the Supplemental Material. Download the .html file and open
it using your internet browser. Hovering over the LAD polygons provides
the mean internet speeds.

\hypertarget{acknowledgement}{%
\section*{Acknowledgement}\label{acknowledgement}}
\addcontentsline{toc}{section}{Acknowledgement}

The authors would like to thank
\url{https://www.broadbandspeedchecker.co.uk/} for providing the main
data used for this paper.

### 2021-04-19 src/online_appendix.Rmd
submission

### Description

The below map illustrates the different clusters of LAD based on the upload 
experienced internet speeds. Hovering over the LAD polygons, the map also presents
the mean AM upload internet speed and the relevant ranking in the country for each 
LAD.

**Is there a chance to briefly describe the clusters with 1-2 words?**

@@

...

>>>>

### Cluster description

@@
